---
title: "Hands-on Exercise 1"
author: "Huo Da"
---

# Getting Started

## Install and Launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)

```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

Note that read_csv can deal with attribute with space in names, while read.csv cannot.

For example, the attribute name is "Project Name". Using read.csv, then the name will be read as "Project.Name". Using read_csv, the name will be read as "Project Name", the space is retained back.

# Data Visualisation

## Plotting No_of_Stu for Different Races

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar()

```

## Plotting Math Grades for students of different gender in each class

```{r}
ggplot(data = exam_data,
       aes(x= CLASS,y=MATHS)) +
  geom_boxplot() +
  facet_grid(~GENDER)

```

```{r}
ggplot(data = exam_data,
       aes(x= CLASS,y=MATHS)) +
  geom_boxplot() +
  facet_grid(rows = vars(GENDER))
```

```{r}
ggplot(data = exam_data,
       aes(x= GENDER,y=MATHS)) +
  geom_boxplot() +
  facet_grid(rows = vars(GENDER),cols = vars(CLASS))+
  theme(axis.text=element_text(size = 6))
```

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar() +
  coord_flip()+
  theme_minimal()+
  theme(panel.background = element_rect(fill = "lightblue", color = "white"))
```

```{r}
ggplot(data = exam_data,
       aes(x = reorder(RACE,RACE,function(x)-length(x)))) +
  geom_bar()+
  labs(y = "No. of Pupils",x = "Race") 
```

```{r}
ggplot(data = exam_data,
       aes(x = fct_infreq(RACE)))+
  geom_text(stat = 'count',aes(label = after_stat(count)),nudge_y = 5)+
  geom_text(stat = 'count',aes(label = scales::percent(after_stat(count/sum(count)))), nudge_y = 15)+
  geom_bar() +
  labs(y = "No. of Pupils",x = "Race") +
  theme(axis.title.y = element_text(hjust = 1)) 
  



```

```{r}
ggplot(data = exam_data,
       aes(x=MATHS))+
  geom_histogram(bins = 20,
                 color = "black",
                 fill = "light blue") +
  geom_vline(xintercept = mean(exam_data$MATHS),
             color = "red",
             linetype = "longdash") +
  geom_vline(xintercept = median(exam_data$MATHS),
             color = "black",
             linetype = "longdash")



```

Reference: https://drsimonj.svbtle.com/plotting-background-data-for-groups-with-ggplot2

```{r}
exam_without_gender = exam_data[,-3]
ggplot(data=exam_data,
       aes(x = ENGLISH, fill = GENDER)) +
  geom_histogram(data = exam_without_gender,fill = "grey",binwidth = 2.5,alpha = 0.5) +
  geom_histogram(binwidth = 2.5,color = "black") +
  facet_wrap(~GENDER) +
  ylim(0,30)+
  guides(fill = FALSE) #remove legend

```

```{r}
ggplot(data=exam_data,
       aes(x = MATHS,y=ENGLISH)) +
  geom_point()+
  xlim(0,100)+
  ylim(0,100)+
  geom_vline(xintercept = 50,linetype = "longdash") +
  geom_hline(yintercept = 50, linetype = "longdash")



```
